---
title: "hw02-seraphinearnold"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework 02: Explore Gapminder and use dplyr

## Bring rectangular data in

```{r}
#install.packages("tidyverse")
#install.packages("gapminder")
```

```{r}
library(gapminder)
library(tidyverse)
```

## Smell test the data

### Is it a data.frame, a matrix, a vector, a list?
### What is its class?
### How many variables/columns?
### How many rows/observations?
 
```{r}
str(gapminder)
```
 
*  I's a data.frame
* Classes are 'tbl_df', 'tbl' and 'data.frame'
  This could also be identified with "class(gapminder)"
* There are 1704 rows/observations of  6 variables/columns

 
### Can you get these facts about “extent” or “size” in more than one way? Can you imagine different functions being useful in different contexts?

Another option would be to use glimpse:
```{r}
glimpse(gapminder)
```

Or specifically ncol and nrow:

```{r}
ncol(gapminder)
nrow(gapminder)
```


### What data type is each variable?

* country  : Factor 
* continent: Factor
* year     : int  
* lifeExp  : num  
* pop      : int  
* gdpPercap: num  

## Explore individual variables

```{r}
summary(gapminder)
```
* Categorial variable I picked: Country
* Quantitative variable I picked: lifeExp


### What are possible values (or range, whichever is appropriate) of each variable?
```{r}
summary(gapminder$country)
```
For a categorial variable like country it makes no sense to look the mean or range, etc. Looking at the levels would make sense though

```{r}
levels(gapminder$country)
```


```{r}
summary(gapminder$lifeExp)
```
For lifeExp it's quite helpful and interesting to explore these values. You could use a quick descriptive statistic with the summary command or specifically explore others like

```{r}
min(gapminder$lifeExp, na.rm = T)
max(gapminder$lifeExp, na.rm = T)
range(gapminder$lifeExp)
```
Or
```{r}
mean(gapminder$lifeExp, na.rm = T)
```
```{r}
median(gapminder$lifeExp, na.rm = T)
```

### What values are typical? What’s the spread? What’s the distribution? Etc., tailored to the variable at hand.
```{r}
sd(gapminder$lifeExp, na.rm = T)
```
```{r}
var(gapminder$lifeExp, na.rm = T)
```


## Explore various plot types

### Relation between lifeExp and country
```{r}
ggplot(gapminder, aes(country, lifeExp))+
  geom_point()+
  xlab("country") +
  ylab("LifeExp") +
  ggtitle("Scatter plot life Expectation and Country")
```


Histogram of LifeExpectancy
```{r}
ggplot(gapminder, aes(lifeExp)) +
    geom_histogram(bins = 50)
```


Life Expectancy Kernel Density Plot
```{r}
  ggplot(gapminder, aes(lifeExp)) +
    geom_histogram(aes(y=..density..)) +
    geom_density()
```


## Use filter(), select() and %>%

Using the select function
```{r}
select(gapminder, lifeExp) %>% 
  summary()
```


LifeExp density only for Europe (using the filter function)

```{r}
gapminder %>% 
  filter(continent == "Europe") %>% 
  ggplot(aes(lifeExp)) +
  geom_density()

```


Boxplot LifeExp only for Europa:

```{r}
gapminder %>% 
  select(continent, lifeExp) %>% 
  filter(continent == 'Europe') %>% 
  ggplot(aes(continent, lifeExp)) + 
  scale_y_log10() +
  geom_boxplot(aes(fill = continent), alpha = 0.2) + 
  geom_jitter(alpha = 0.2)
```



## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
